Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-08 paperIEEE.tex
minor fixes
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@119 db7c848b-7755-4df2-b763-8796797afa52

We conducted performance benchmarks for the decomposition modes on a
rendering cluster using hexacore Intel Xeon E5-2620v3 CPUs (2.4\,GHz), 16\,GB DDR4 RAM per node, nVidia GTX 970 GPUs with 4\,GB VRAM each, and QDR Infiniband.
GCC 4.8 has been used with CMake 3.7 release mode settings to compile the software stack.

We tested the decomposition modes with both volume data (\fig{fBenchmarkEVolve})
and polygonal data  (\fig{fBenchmarkEqPly}) and used test scenes that allowed us
to finely adjust the rendering load the system has to cope with. In both cases
the scene comprised of two rows of instantiated, identical models with 30 models
in total. Rendering was performed at a resolution of 2560x1440. The camera is
initially placed in the center of the scene, between the two rows
(\fig{fBenchmarkEVolve} and \fig{fBenchmarkEqPly}), rendering only half of the
model instances. It is then moved backward over the duration of 800 frames,
steadily increasing the rendering load by revealing more models, until all 30
instances are visible.

We investigated the scalability of individual decomposition modes by running the
same experiment using a varying number of render nodes (2-9), as well as one
dedicated application/display node. We subsequently summed up the duration of
all rendered frames for each run (\fig{fCompounds}). We provide the results of
the load balancer providing the better result for each application and compound.
The simpler \textsf{tree\_equalizer} outperforms in almost all cases the
load-grid-driven \textsf{load\_equalizer}, except for sort-first volume
rendering where the load in the region of interest is relatively uniform. This
counterintuitive result once more proves that simple algorithms often outperform
theoretically better, but more complex implementations.

@@

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Flower volume model at $1024^3$ unsigned byte (1\,GB).}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} David statue at 1\,mm resolution (988\,MB).}
\end{figure}

@@

The data distribution benchmarks have been performed on a cluster using
dual processor Intel Xeon X5690 CPUs (3.47\,GHz), 24\,GB main memory per node, 10\,GBit/s
Ethernet and QDR Infiniband. Intel ICC 2017 has been used with CMake 3.2 release
mode settings to compile the software stack.

@@

The volume data sets are distributed in a single object, serializing the raw
volume buffer. The Spike volume data set has a significant compression ratio,
which is reflected by the results in \fig{fSpikesDist}. Compression for this
data is beneficial for transmitting data over a 10\,Gb/s %why was there a tilde inbetween? Should this be ~10Gb/s instead?
link even for a single
slave process. Buffering has little benefit since the serialization of volume
data is trivial. Buffered compression makes a significant difference, since the
compression cost can be amortized over $n$ nodes, reaching raw data transmission
rates of 3.7\,GB/s with the default Snappy compressor and at best 4.4\,GB/s with
ZStandard at level 1.

>>>>

We conducted new performance benchmarks for various decomposition modes on a cluster
using hexacore Intel Xeon E5-2620v3 CPUs (2.4\,GHz), nVidia GTX 970 GPUs with 4\,GB VRAM each, 16\,GB main
memory per node, 4\,GBit/s Ethernet, and QDR Infiniband. GCC 4.8 has been used with CMake
3.7 release mode settings to compile the software stack.

We tested the decomposition modes with both volume data (\fig{fBenchmarkEVolve})
and polygonal data  (\fig{fBenchmarkEqPly}) and used test scenes that allowed us
to continuously modify the rendering load the system has to cope with. In both cases
the scene comprised of two rows of instantiated, identical models with 30 models
in total. Rendering was performed at an output resolution of $2560\times1440$. The camera is
initially placed in the center of the scene, between the two rows
(\fig{fBenchmarkEVolve} and \fig{fBenchmarkEqPly}), rendering only half of the
model instances. It is then moved backward over the duration of 800 frames,
steadily increasing the rendering load by revealing more models, until all 30
instances are visible.

We investigated the scalability of individual decomposition modes by running the
same experiment using a varying number of render nodes (2-9), as well as one
dedicated application/display node. We subsequently summed up the duration of
all rendered frames for each run (\fig{fCompounds}), backing up along the alley of models.
We provide the results of the load balancer providing the better result for each application and compound.\FIXME{what does this last sentence mean?}
The simpler \textsf{tree\_equalizer} outperforms in almost all cases the
load-grid-driven \textsf{load\_equalizer}, except for sort-first volume
rendering where the load in the region of interest is relatively uniform. This
counterintuitive result once more proves that simple algorithms often outperform
theoretically better, but more complex implementations.

\FIXME{oscillation, pixel, clearer}
\FIXME{I'm not sure I read Fig.10 the same way. Left it seems that tile\_equalizer is the best for polygonal rendering, it's constantly faster than the tree\_equalizer mentioned in the paragraph above.}

@@

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eVolve}
  \caption{\label{fBenchmarkEVolve} Flower alley with 30 volume models of $1024^3$ unsigned bytes (1\,GB) each.}
\end{figure}

\begin{figure}[ht]\center
  \includegraphics[width=.9\columnwidth]{images/benchmark_eqPly2}
  \caption{\label{fBenchmarkEqPly} David alley with 30 statues of 56m triangles (988\,MB) each.\FIXME{check the triangle count}}
\end{figure}

@@

The data distribution benchmarks have been performed on a cluster using
dual processor Intel Xeon X5690 CPUs (3.47\,GHz), 24\,GB main memory, 10\,GBit/s
Ethernet and QDR Infiniband. Intel ICC 2017 has been used with CMake 3.2 release
mode settings to compile the software stack.

@@

The volume data sets are distributed in a single object, serializing the raw
volume buffer. The Spike volume data set has a significant compression ratio,
which is reflected by the results in \fig{fSpikesDist}. Compression for this
data is beneficial for transmitting data over a 10~Gb/s \FIXME{~10\,Gb/s ?} link even for a single
slave process. Buffering has little benefit since the serialization of volume
data is trivial. Buffered compression makes a significant difference, since the
compression cost can be amortized over $n$ nodes, reaching raw data transmission
rates of 3.7~GB/s \FIXME{~3.7\,GB/s ?} with the default Snappy compressor and at best 4.4~GB/s \FIXME{~4.4\,GB/s ?} with
ZStandard at level 1.
